---
title: "Week 4 - Homework"
author: "STAT 420, Summer 2020, Xuenan Mi, NetID: xmi4"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---


***

## Exercise 1 (Using `lm`)

For this exercise we will use the data stored in [`nutrition-2018.csv`](nutrition-2018.csv). It contains the nutritional values per serving size for a large variety of foods as calculated by the USDA in 2018. It is a cleaned version totaling 5956 observations and is current as of April 2018.

The variables in the dataset are:

- `ID` 
- `Desc` - short description of food
- `Water` - in grams
- `Calories` 
- `Protein` - in grams
- `Fat` - in grams
- `Carbs` - carbohydrates, in grams
- `Fiber` - in grams
- `Sugar` - in grams
- `Calcium` - in milligrams
- `Potassium` - in milligrams
- `Sodium` - in milligrams
- `VitaminC` - vitamin C, in milligrams
- `Chol` - cholesterol, in milligrams
- `Portion` - description of standard serving size used in analysis

**(a)** Fit the following multiple linear regression model in `R`. Use `Calories` as the response and `Fat`, `Sugar`, and `Sodium` as predictors.

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \epsilon_i.
\]

Here,

- $Y_i$ is `Calories`.
- $x_{i1}$ is `Fat`.
- $x_{i2}$ is `Sugar`.
- $x_{i3}$ is `Sodium`.

Use an $F$-test to test the significance of the regression. Report the following:
 
- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
nutrition = read.csv("nutrition-2018.csv", header = T, stringsAsFactors = F)
model1 = lm (Calories ~ Fat + Sugar + Sodium, data = nutrition)
summary(model1)
```


- The null hypothesis: $\beta_1 = \beta_2 = \beta_3 = 0$; The alternative hypothesis: At least one of $\beta_1 , \beta_2, \beta_3 \neq 0$
- Test statistics: `r summary(model1)$fstatistic[1]`
- p-value $< 2.2 * 10^{-16}$
- Decision: Reject $H_0$ at $\alpha = 0.01$
- Conclusion: There are some linear relationship between Calories and Fat, Sugar, Sodium.


**(b)** Output only the estimated regression coefficients. Interpret all $\hat{\beta}_j$ coefficients in the context of the problem.

```{r}
summary(model1)$coefficients
```

- $\hat{\beta_0}$ = `r coef(model1)[1]` is the estimated calories of food with 0g fat, 0g sugar and 0 milligram sodium.
- $\hat{\beta_1}$ = `r coef(model1)[2]` is the estimated average change in calories for 1g increase in fat of food with certain sugar and sodium.
- $\hat{\beta_2}$ = `r coef(model1)[3]` is the estimated average change in calories for 1g increase in sugar of food with certain fat and sodium.
- $\hat{\beta_3}$ = `r coef(model1)[4]`is the estimated average change in calories for 1 milligram increase in sodium of food with certain fat and sugar.


**(c)** Use your model to predict the number of `Calories` in a Big Mac. According to [McDonald's publicized nutrition facts](https://www.mcdonalds.com/us/en-us/about-our-food/nutrition-calculator.html), the Big Mac contains 30g of fat, 9g of sugar, and 1010mg of sodium.

```{r}
new_Mac = data.frame(Fat = 30, Sugar = 9, Sodium = 1010)
predict(model1, newdata = new_Mac)
```

Based on the model, the Calories in a Big Mac is 396.2863.

**(d)** Calculate the standard deviation, $s_y$, for the observed values in the Calories variable. Report the value of $s_e$ from your multiple regression model. Interpret both estimates in the context of this problem.

```{r}
(s_y = sd(nutrition$Calories))
(s_e = summary(model1)$sigma)
```

- $s_y$ = `r sd(nutrition$Calories)` is varibility of calories value without any predictors，which indicates how calories vary about the mean.
- $s_e$ = `r summary(model1)$sigma` is the estimated varibility of the residuals of the fitted model, which indicates how calories vary about the fitted value from the model.


**(e)** Report the value of $R^2$ for the model. Interpret its meaning in the context of the problem.

```{r}
summary(model1)$r.squared
```

$R^2$ = `r summary(model1)$r.squared` means that 76.86% of variation in calories can be explained by the linear relationship with fat, sugar and sodium. 


**(f)** Calculate a 90% confidence interval for $\beta_2$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(model1, parm = "Sugar", level = 0.9)
```

The 90% confidence interval for $\beta_2$ is $$(3.783051, 4.017983)$$.

Interpretation: We are 90% confident that the average change in calories for 1g increase in sugar of food with certain fat and sodium is between 3.783051 and 4.017983.


**(g)** Calculate a 95% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(model1, parm = "(Intercept)", level = 0.95)
```

The 95% confidence interval for $\beta_0$ is $$(97.69443, 103.2177)$$.

Interpretation: We are 95% confident that the mean calories of food with with 0g fat, 0g sugar and 0 milligram sodium is between 97.69443 and 103.2177.


**(h)** Use a 99% confidence interval to estimate the mean Calorie content of a food with 23g of fat, 0g of sugar, and 400mg of sodium, which is true of a large order of McDonald's french fries. Interpret the interval in context.

```{r}
new_food = data.frame(Fat = 23, Sugar = 0, Sodium = 400)
predict(model1, newdata = new_food, interval = "confidence", level = 0.99)
```

We are 99% confident that the mean calories for food with 23g fat, 0g sugar, and 400mg sodium is in the interval $$(294.3532, 301.7224)$$

 
**(i)** Use a 99% prediction interval to predict the Calorie content of a Crunchwrap Supreme, which has 21g of fat, 6g of sugar, and 1200mg of sodium according to [Taco Bell's publicized nutrition information](https://www.tacobell.com/nutrition/info). Interpret the interval in context.

```{r}
new_food = data.frame(Fat = 21, Sugar = 6, Sodium = 1200)
predict(model1, newdata = new_food, interval = "prediction", level = 0.99)
```

We are 99% confident that the calories for food with 21g fat, 6g sugar, and 1200mg sodium is in the interval $$(101.0345, 517.7786)$$


***

## Exercise 2 (More `lm` for Multiple Regression)

For this exercise we will use the data stored in [`goalies.csv`](goalies.csv). It contains career data for 462 players in the National Hockey League who played goaltender at some point up to and including the 2014-2015 season. The variables in the dataset are:
 
- `W` - Wins
- `GA` - Goals Against
- `SA` - Shots Against
- `SV` - Saves
- `SV_PCT` - Save Percentage
- `GAA` - Goals Against Average
- `SO` - Shutouts
- `MIN` - Minutes
- `PIM` - Penalties in Minutes

For this exercise we will consider three models, each with Wins as the response. The predictors for these models are:

- Model 1: Goals Against, Saves
- Model 2: Goals Against, Saves, Shots Against, Minutes, Shutouts
- Model 3: All Available

**(a)** Use an $F$-test to compares Models 1 and 2. Report the following:

- The null hypothesis
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- The model you prefer

```{r}
goalies = read.csv("goalies.csv", header = T, stringsAsFactors = F)
Model1 = lm(W ~ GA + SV, data = goalies) 
Model2 = lm(W ~ GA + SV + SA + MIN + SO, data = goalies)
anova(Model1, Model2)
```

- $H_0$: $\beta_{SA}=\beta_{MIN}=\beta_{SO}=0$
- $H_1$: At least one of $\beta_{SA}, \beta_{MIN}, \beta_{SO} \neq 0$
- Test statistic: $462.59$
- p value $< 2.2*10^{-16}$
- Decision: Reject $H_0$ at $\alpha = 0.05$
- Model prefer: We will prefer Model 2 with predictor: Goals Against, Saves, Shots Against, Minutes, Shutouts.


**(b)** Use an $F$-test to compare Model 3 to your preferred model from part **(a)**. Report the following:

- The null hypothesis
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- The model you prefer

```{r}
Model3 = lm (W ~., data = goalies)
anova(Model2, Model3)
```

- $H_0$: $\beta_{SV-PCT}=\beta_{GAA}=\beta_{PIM}=0$
- $H_1$: At least one of $\beta_{SV-PCT}, \beta_{GAA}, \beta_{PIM} \neq 0$
- Test statistic: $4.052$
- p value: $0.007353$
- Decision: Reject $H_0$ at $\alpha = 0.05$
- Model prefer: We will prefer Model 3 with all predictors.


**(c)** Use a $t$-test to test $H_0: \beta_{\texttt{SV}} = 0 \ \text{vs} \ H_1: \beta_{\texttt{SV}} \neq 0$ for the model you preferred in part **(b)**. Report the following:

- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$

```{r}
summary(Model3)$coefficients['SV', 't value']
summary(Model3)$coefficients['SV', 'Pr(>|t|)']
```

- Test statistic: $-3.857739$
- p value: $0.0001310371$
- Decision: Reject $H_0$ at $\alpha = 0.05$


***

## Exercise 3 (Regression without `lm`)

For this exercise we will once again use the `Ozone` data from the `mlbench` package. The goal of this exercise is to fit a model with `ozone` as the response and the remaining variables as predictors.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Obtain the estimated regression coefficients **without** the use of `lm()` or any other built-in functions for regression. That is, you should use only matrix operations. Store the results in a vector `beta_hat_no_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_no_lm ^ 2)`.

```{r}
n = nrow(Ozone)
X = as.matrix(cbind(rep(1, n), Ozone[,-1]))
y = Ozone$ozone
(beta_hat_no_lm = as.vector(solve(t(X) %*% X) %*% t(X) %*% y))
sum(beta_hat_no_lm ^ 2)
```


**(b)** Obtain the estimated regression coefficients **with** the use of `lm()`. Store the results in a vector `beta_hat_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_lm ^ 2)`.

```{r}
model3b = lm(ozone ~., data = Ozone)
(beta_hat_lm = as.vector(coef(model3b)))
sum(beta_hat_lm ^ 2)
```


**(c)** Use the `all.equal()` function to verify that the results are the same. You may need to remove the names of one of the vectors. The `as.vector()` function will do this as a side effect, or you can directly use `unname()`.

```{r}
all.equal(beta_hat_no_lm, beta_hat_lm)
```


**(d)** Calculate $s_e$ without the use of `lm()`. That is, continue with your results from **(a)** and perform additional matrix operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.

```{r}
y_hat = X %*% beta_hat_no_lm
e = y - y_hat
(Se_no_lm = sqrt( t(e) %*% e / (n - 4)))
(Se_lm = summary(model3b)$sigma)
Se_no_lm == Se_lm
```

Without the use of `lm()`, $s_e$ = 4.806115; Based on the results from `lm()`, $s_e$ = 4.806115. They are same.


**(e)** Calculate $R^2$ without the use of `lm()`. That is, continue with your results from **(a)** and 
**(d)**, and perform additional operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.

```{r}
SST = sum((y - mean (y)) ^ 2)
SSR = sum((y_hat - mean(y)) ^ 2)
(R2_no_lm = SSR / SST)
(R2_lm = summary(model3b)$r.squared)

```

Without the use of `lm()`, $R^2$ = 0.6398887; Based on the results from `lm()`, $R^2$ = 0.6398887. They are same.


***

## Exercise 4 (Regression for Prediction)

For this exercise use the `Auto` dataset from the `ISLR` package. Use `?Auto` to learn about the dataset. The goal of this exercise is to find a model that is useful for **predicting** the response `mpg`. We remove the `name` variable as it is not useful for this analysis. (Also, this is an easier to load version of data from the textbook.)

```{r}
# load required package, remove "name" variable
library(ISLR)
Auto = subset(Auto, select = -c(name))
```

When evaluating a model for prediction, we often look at RMSE. However, if we both fit the model with all the data as well as evaluate RMSE using all the data, we're essentially cheating. We'd like to use RMSE as a measure of how well the model will predict on *unseen* data. If you haven't already noticed, the way we had been using RMSE resulted in RMSE decreasing as models became larger.

To correct for this, we will only use a portion of the data to fit the model, and then we will use leftover data to evaluate the model. We will call these datasets **train** (for fitting) and **test** (for evaluating). The definition of RMSE will stay the same

\[
\text{RMSE}(\text{model, data}) = \sqrt{\frac{1}{n} \sum_{i = 1}^{n}(y_i - \hat{y}_i)^2}
\]

where

- $y_i$ are the actual values of the response for the given data.
- $\hat{y}_i$ are the predicted values using the fitted model and the predictors from the data.

However, we will now evaluate it on both the **train** set and the **test** set separately. So each model you fit will have a **train** RMSE and a **test** RMSE. When calculating **test** RMSE, the predicted values will be found by predicting the response using the **test** data with the model fit using the **train** data. *__Test__ data should never be used to fit a model.*

- Train RMSE: Model fit with *train* data. Evaluate on **train** data.
- Test RMSE: Model fit with *train* data. Evaluate on **test** data.

Set a seed of `11`, and then split the `Auto` data into two datasets, one called `auto_trn` and one called `auto_tst`. The `auto_trn` data frame should contain 292 randomly chosen observations. The `auto_tst` data will contain the remaining observations. Hint: consider the following code:

```{r}
set.seed(11)
auto_trn_idx = sample(1:nrow(Auto), 292)
auto_trn = Auto[auto_trn_idx,]
auto_tst = Auto[-auto_trn_idx,]
```

Fit a total of five models using the training data.

- One must use all possible predictors.
- One must use only `displacement` as a predictor.
- The remaining three you can pick to be anything you like. One of these should be the *best* of the five for predicting the response.

For each model report the **train** and **test** RMSE. Arrange your results in a well-formatted markdown table. Argue that one of your models is the best for predicting the response.

```{r}
library(knitr)
#model1 with all predictors
mod1 = lm(mpg ~ ., data = auto_trn)
RMSE_trn_1 = sqrt(mean(mod1$residuals ^ 2))
fitted1 = predict(mod1, newdata = auto_tst[,-1])
RMSE_tst_1 = sqrt(mean((auto_tst$mpg - fitted1) ^ 2))

#model2 with displacement
mod2 = lm(mpg ~ displacement, data = auto_trn)
RMSE_trn_2 = sqrt(mean(mod2$residuals ^ 2))
fitted2 = predict(mod2, newdata = data.frame(displacement = auto_tst$displacement ))
RMSE_tst_2 = sqrt(mean((auto_tst$mpg - fitted2) ^ 2))

#model3 with weight and year
mod3 = lm(mpg ~ weight + year, data = auto_trn)
RMSE_trn_3 = sqrt(mean(mod3$residuals ^ 2))
fitted3 = predict(mod3, newdata = data.frame(auto_tst[,c("weight",'year')]))
RMSE_tst_3 = sqrt(mean((auto_tst$mpg - fitted3) ^ 2))

#model4 with weight and acceleration
mod4 = lm(mpg ~ weight + acceleration , data = auto_trn)
RMSE_trn_4 = sqrt(mean(mod4$residuals ^ 2))
fitted4 = predict(mod4, newdata = data.frame(auto_tst[,c("weight",'acceleration')]))
RMSE_tst_4 = sqrt(mean((auto_tst$mpg - fitted4) ^ 2))

#model5 with displacement and year
mod5 = lm(mpg ~ displacement  + year , data = auto_trn)
RMSE_trn_5 = sqrt(mean(mod5$residuals ^ 2))
fitted5 = predict(mod5, newdata = data.frame(auto_tst[,c("displacement",'year')]))
RMSE_tst_5 = sqrt(mean((auto_tst$mpg - fitted5) ^ 2))


#########

df = data.frame(
  model = c('model1', 'model2', 'model3', 'model4', 'model5'),
  train_RMSE = c(RMSE_trn_1, RMSE_trn_2, RMSE_trn_3, RMSE_trn_4, RMSE_trn_5),
  test_RMSE = c(RMSE_tst_1, RMSE_tst_2, RMSE_tst_3, RMSE_tst_4, RMSE_tst_5)
)

kable(df)
```

Based on the RMSE of five models, the model3 has the smallest test RMSE. So the model with weight and year as predictors is the best for predicting the response.

***

## Exercise 5 (Simulating Multiple Regression)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \beta_4 x_{i4} + \beta_5 x_{i5} + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 2$
- $\beta_1 = -0.75$
- $\beta_2 = 1.5$
- $\beta_3 = 0$
- $\beta_4 = 0$
- $\beta_5 = 2$
- $\sigma^2 = 25$

We will use samples of size `n = 42`.

We will verify the distribution of $\hat{\beta}_2$ as well as investigate some hypothesis tests.

**(a)** We will first generate the $X$ matrix and data frame that will be used throughout the exercise. Create the following nine variables:

- `x0`: a vector of length `n` that contains all `1`
- `x1`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `x2`: a vector of length `n` that is randomly drawn from a uniform distribution between `0` and `4`
- `x3`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `1`
- `x4`: a vector of length `n` that is randomly drawn from a uniform distribution between `-2` and `2`
- `x5`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `X`: a matrix that contains `x0`, `x1`, `x2`, `x3`, `x4`, and `x5` as its columns
- `C`: the $C$ matrix that is defined as $(X^\top X)^{-1}$
- `y`: a vector of length `n` that contains all `0`
- `sim_data`: a data frame that stores `y` and the **five** *predictor* variables. `y` is currently a placeholder that we will update during the simulation.

Report the sum of the diagonal of `C` as well as the 5th row of `sim_data`. For this exercise we will use the seed `420`. Generate the above variables in the order listed after running the code below to set a seed.

```{r}
set.seed(420)
sample_size = 42
```

```{r}
x0 = rep(1, sample_size)
x1 = rnorm(sample_size, 0, 2)
x2 = runif(sample_size, 0, 4)
x3 = rnorm(sample_size, 0, 1)
x4 = runif(sample_size, -2, 2)
x5 = rnorm(sample_size, 0, 2)
X = cbind(x0, x1, x2, x3, x4, x5)
C = solve(t(X) %*% X)
y = rep(0, sample_size)
sim_data = data.frame(cbind(y, x1, x2, x3, x4, x5))
sum(diag(C))
sim_data[5, ]
```


**(b)** Create three vectors of length `2500` that will store results from the simulation in part **(c)**. Call them `beta_hat_1`, `beta_3_pval`, and `beta_5_pval`.

```{r}
sim_num = 2500
beta_hat_1 = rep(0, sim_num)
beta_3_pval = rep(0, sim_num)
beta_5_pval = rep(0, sim_num)

```


**(c)** Simulate 2500 samples of size `n = 42` from the model above. Each time update the `y` value of `sim_data`. Then use `lm()` to fit a multiple regression model. Each time store:

- The value of $\hat{\beta}_1$ in `beta_hat_1`
- The p-value for the two-sided test of $\beta_3 = 0$ in `beta_3_pval`
- The p-value for the two-sided test of $\beta_5 = 0$ in `beta_5_pval`

```{r}
beta_0 = 2
beta_1 = -0.75
beta_2 = 1.5
beta_3 = 0
beta_4 = 0
beta_5 = 2
sigma = 5

for (i in 1:sim_num){
  eps = rnorm(sample_size, 0, sigma)
  sim_data$y = beta_0 * x0 + beta_1 * x1 + beta_2 * x2 + beta_3 * x3 + beta_4 * x4 + beta_5 * x5 + eps
  fit = lm(y ~ x1 + x2 + x3 + x4 + x5, data = sim_data)
  beta_hat_1[i] = coef(fit)[2]
  beta_3_pval[i] = summary(fit)$coefficients['x3', 'Pr(>|t|)']
  beta_5_pval[i] = summary(fit)$coefficients['x5', 'Pr(>|t|)']
}
```


**(d)** Based on the known values of $X$, what is the true distribution of $\hat{\beta}_1$?

\[
  \hat{\beta}_1\sim N\left(\beta_1, \sigma^2C_{11}\right)
\]

```{r}
(mean = beta_1)
(var = sigma ^ 2 * C[2,2])
```

So the true distribution of $\hat{\beta}_1$ is
\[
  \hat{\beta}_1\sim N\left(-0.75, 0.1892102\right)
\]


**(e)** Calculate the mean and variance of `beta_hat_1`. Are they close to what we would expect? Plot a histogram of `beta_hat_1`. Add a curve for the true distribution of $\hat{\beta}_1$. Does the curve seem to match the histogram?

```{r}
mean(beta_hat_1)
var(beta_hat_1)
```

The results from simulation are close to true distribution.


```{r}
hist(beta_hat_1, prob = T, breaks = 20, xlab = expression(hat(beta)[1]), main = '', border = 'blue')
curve(dnorm(x, mean = beta_1,sd = sqrt(sigma ^ 2 * C[2,2])), col = 'darkorange', add = T, lwd = 3)
```

The curve of true distribution matches the histogram of simulation data.


**(f)** What proportion of the p-values stored in `beta_3_pval` is less than 0.10? Is this what you would expect?

```{r}
mean (beta_3_pval < 0.10)
```

Because $\beta_3=0$, we expect about 10% of the p values to be significant at $\alpha = 0.1$ by chance. So it matches our expectation.


**(g)** What proportion of the p-values stored in `beta_5_pval` is less than 0.01? Is this what you would expect?

```{r}
mean (beta_5_pval < 0.01)
```

Because $\beta_5 \neq 0$, we expect most of the p values to be significant at $\alpha = 0.01$. So it matches our expectation.

